### 2019-12-13 thesis/txt/1-related-work.tex
adding various examples to rw, data set and study

\subsection{Continuous Integration}
The first paragraphs of this section present the results of past research into CI: its influence on the software engineering processes, the reasons for CI builds to fail and the role of testing and static analysis in CI builds.
Later we describe how these researchers obtain the data sets these works are based on.
We show how they analyze build logs and why the chunk retrieval techniques we investigate can support them to broaden their studies.
% This section starts with an overview on research into why CI is used and its impact on development.
% It presents analyses on why CI builds fail and how testing and static analysis tools are used within CI\@.
% We explain how researchers compose data sets for their work on CI and how chunk retrieval techniques can support them.

@@

\subsection{System Log Analysis}
\label{sec:log-analysis}
A related field of log processing is the processing of system log files produced during runtime.
A main difference between build logs and system logs is that system logs are fundamentally structured through events.
Each line in the log file represents one event.
The first goal in parsing system log files is to separate constant and variable parts within a log message~\cite{nagappan2010abstracting,he2017towards}.
Next, the log messages are clustered into log events, unifying messages with identical constant parts and varying parameters.
The output of a log parser is a structured log, composed of a list of timed events and the corresponding parameter values~\cite{he2016evaluation}.
This structured log is then the input to various machine learning and data mining processes.
Researchers mine patterns for operational profiling~\cite{nagappan2009efficiently}, debugging~\cite{oliner2012advances}, performance analytics or anomaly detection~\cite{nagappan2010abstracting}.
Xu et al.~\cite{xu2009detecting} leverage the connection of log statements to the source code producing them to separate messages into constant and variable parts more accurately.

@@

@@

The text instantiation of FlashExtract models the extraction of a single substring as a pair of two cut positions.
A position is either specified by an absolute character index or by a pair of two regular expressions.
The first regular expression matches the substring directly before the characterized position, the second regular expression matches the substring directly after.
A regular expression in FlashExtract is a concatenation of tokens, e.g.\ standard character classes or string literals frequently occurring in the input examples.

@@

>>>>

\subsection{Continuous Integration}
\label{sec:rw-ci}
The first paragraphs of this section present the results of past research into CI: its influence on the software engineering processes, the reasons for CI builds to fail and the role of testing and static analysis in CI builds.
Later we describe how these researchers obtain the data sets these works are based on.
We show how they analyze build logs and why the chunk retrieval techniques we investigate can support them to broaden their studies.
% This section starts with an overview on research into why CI is used and its impact on development.
% It presents analyses on why CI builds fail and how testing and static analysis tools are used within CI\@.
% We explain how researchers compose data sets for their work on CI and how chunk retrieval techniques can support them.

@@

\subsection{System Log Analysis}
\label{sec:log-analysis}
A related field of log processing is the processing of system log files produced during runtime.
A main difference between build logs and system logs is that system logs are fundamentally structured through events.
Each line in the log file represents one event with a set of fields: timestamp, verbosity level and raw message content~\cite{he2017towards}.
Listing \ref{lst:system-log} shows example lines from a system log.

The first goal in parsing system log files is to separate constant and variable parts within a log message~\cite{nagappan2010abstracting,he2017towards}.
Next, the log messages are clustered into log events, unifying messages with identical constant parts and varying parameters.
The output of a log parser is a structured log, composed of a list of timed events and the corresponding parameter values~\cite{he2016evaluation}.
This structured log is then the input to various machine learning and data mining processes.
Researchers mine patterns for operational profiling~\cite{nagappan2009efficiently}, debugging~\cite{oliner2012advances}, performance analytics or anomaly detection~\cite{nagappan2010abstracting}.
Xu et al.~\cite{xu2009detecting} leverage the connection of log statements to the source code producing them to separate messages into constant and variable parts more accurately.

@@

\lstset{
  morekeywords={INFO, WARN, 2008-11-09},
  postbreak=\mbox{\textcolor{cyan}{$\hookrightarrow$}\space},
  showspaces=false,
  showstringspaces=false,
  keywordstyle=\color{blue},
	frame=single,
  extendedchars=false,
  texcl=false
}
\begin{figure}[]
  \centering
  \begin{lstlisting}[breaklines=true]
2008-11-09 20:46:55,556 INFO dfs.DataNode$Packet- Responder: Received block blk_3587508140051953248 of size 67108864 from /10.251.42.84
2008-11-09 20:49:46,764 WARN PacketResponder 0 for block blk\_3587508140051953248 terminating
  \end{lstlisting}  
  \caption{System Log Statements, adapted example from~\cite{he2017towards}}
  \label{lst:system-log}
\end{figure}

@@

The text instantiation of FlashExtract models the extraction of a single substring as a pair of two cut positions.
A position is either specified by an absolute character index or by a pair of two regular expressions.
The first regular expression matches the substring directly before the characterized position, the second regular expression matches the substring directly after.
A regular expression in FlashExtract is a concatenation of tokens, e.g.\ standard character classes or string literals frequently occurring in the input examples.
Listing \ref{lst:prose-program} shows a text extraction program synthesized by FlashExtract.
This program defines the first position as after a colon followed by a newline character and before a piece of text with all capital letters.
It defines the second position as before two newline characters.

@@

\lstset{
  language=caml,
  morekeywords={PosToEndRegion, RegexPosition, RegexPair, StartToPosRegion},
  postbreak=\mbox{\textcolor{cyan}{$\hookrightarrow$}\space},
  showspaces=false,
  showstringspaces=false,
  stringstyle=\color{blue},
  keywordstyle=\bfseries\color{black},
	frame=single,
  extendedchars=false,
  escapeinside=//
}
\begin{figure}[]
  \centering
  \begin{lstlisting}[breaklines=true]
let s = v in let s = PosToEndRegion(s, RegexPosition(s, RegexPair("Colon/$\circ$/Line Separator", "ALL CAPS"), 1)) in StartToPosRegion(s, RegexPosition(s, RegexPair("/$\varepsilon$/", "Line Separator/{\color{blue}$\circ$}/Line Separator"), 1))
  \end{lstlisting}  
  \caption{Text extraction program synthesized by FlashExtract}
  \label{lst:prose-program}
\end{figure}

### 2019-12-13 thesis/txt/2-models.tex
adding various examples to rw, data set and study

\begin{table}[htbp]
\centering
\caption{Chunk Retrieval Techniques}
\begin{tabularx}{\textwidth}{@{}XlXlX@{}} 
\toprule
Name                         & Acronym & Identification Technique                                   & Granularity & Configuration             \\ 
\midrule
Program Synthesis by Example & PBE     & Regular expression program                                 & Character   & In / Output Examples      \\
Common Text Similarity       & CTS     & Tf-idf \& cosine similarity, mean number of lines expected & Line        & Output Examples           \\
Keyword Search               & KWS     & Keywords, mean number of lines expected                    & Line        & Keywords, Context Length  \\
Random Line Retrieval        & RLR     & Random sample                                              & Line        & Retrieval Length          \\
Diff Approach                &         & Line not present in successful log, information retrieval  & Line        & Successful Log File       \\
\bottomrule
\end{tabularx}
\label{tab:ctr}
\end{table}

@@

\subsection{Common Text Similarity (CTS)}
\label{sec:expl-ts}
Text Similarity approaches are used to filter unstructured textual software artifacts~\cite{runeson2007detection,marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
One common and simple technique is the Vector Space Model~\cite{schutze2008introduction}.
We investigate when text similarity is a suitable technique to retrieve information chunks from build logs.
In the following we will explain the concept of how we apply text similarity to information retrieval from CI build logs (CTS), the concrete implementation is described in~\ref{sec:impl-ts}.

@@

\subsection{Keyword Search (KWS)}
\label{sec:expl-skws}
When developers are looking for a specific piece of information within a large amount of unstructured information, a first ad-hoc approach they use is searching for related keywords.
This was one of the most common approaches we took when searching for the reason the build failed within a log while creating our \emph{LogChunks} data set.
As this is a technique readily available in many tools developers use to view build logs, we study when such a keyword search is suitable for retrieving information chunks from CI build logs.
In the following we will explain how we use simple keyword search to retrieve information from CI build logs (KWS).

>>>>

\begin{table}[htbp]
\centering
\caption{Chunk Retrieval Techniques}
\begin{tabularx}{\textwidth}{@{}XlXlX@{}} 
\toprule
Name                         & Acronym & Identification Technique                                   & Granularity & Configuration             \\ 
\midrule
Program Synthesis by Example & PBE     & Regular expression program                                 & Character   & In / output examples      \\
Common Text Similarity       & CTS     & Tf-idf \& cosine similarity, mean number of lines expected & Line        & Output examples           \\
Keyword Search               & KWS     & Keywords, mean number of lines expected                    & Line        & Keywords, context length  \\
Random Line Retrieval        & RLR     & Random sample                                              & Line        & Retrieval length          \\
Diff Approach                &         & Line not present in successful log, information retrieval  & Line        & Logs from failing and successful builds      \\
\bottomrule
\end{tabularx}
\label{tab:ctr}
\end{table}

@@

\subsection{Common Text Similarity (CTS)}
\label{sec:expl-ts}
Text Similarity approaches are used to filter unstructured textual software artifacts~\cite{runeson2007detection,marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
One common and simple technique is the Vector Space Model~\cite{schutze2008introduction}.
We investigate when text similarity is a suitable technique to retrieve information chunks from build logs.
In the following we will explain the concept of how we apply text similarity to information retrieval from CI build logs, which we refer to as \emph{CTS}\@.

@@

\subsection{Keyword Search (KWS)}
\label{sec:expl-skws}
When developers are looking for a specific piece of information within a large amount of unstructured information, a first ad-hoc approach they use is searching for related keywords.
This was one of the most common approaches we took when searching for the reason the build failed within a log while creating our \emph{LogChunks} data set.
As this is a technique readily available in many tools developers use to view build logs, we study when such a keyword search is suitable for retrieving information chunks from CI build logs.
In the following we will explain how we use simple keyword search to retrieve information from CI build logs, which we refer to as \emph{KWS}\@.

### 2019-12-13 thesis/txt/3-data-set.tex
adding various examples to rw, data set and study

\chapter{LogChunks Data Set}
\label{sec:data-set}
\includegraphics[page=5, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
This chapter describes the creation of the \emph{LogChunks} data set that we use in our empirical comparison of chunk retrieval techniques.
\emph{LogChunks} is a collection of 797 Travis CI build logs from 80 GitHub repositories spread over 29 programming languages.
For each build log we manually labeled which substring describes why the build failed.
The data set also provides keywords a developer would use to search for the labeled log chunk and categorizes the log chunks according to their format within the log.

@@

\section{Motivation}
\todo{In/output examples (I/O examples) are used to configure two of the information retrieval techniques we are evaluating for this thesis, PROSE regular expression program synthesis by example and text similarity.
One example set always contains build log in/output examples from the same software repository or project and for one specific build log information. They are all from one repository because we investigate information retrieval techniques which are always configured to the scope of one particular software repository or project.
Our intuition was that the ability of PROSE to be able to successfully learn a regex program depended on the structural uniformity of the provided in/output examples.
The regular expressions need a consistent pattern in the build log string at the borders or around the textual representation of a build log information to match on for the retrieval.
To be able to quantify this intuition in our evaluation we assigned \emph{structural categories} to each of the examples within an example set.}

@@

\lstset{
  language=XML,
  morekeywords={Examples, Example, Log, Keywords, Category, Chunk},
  postbreak=\mbox{\textcolor{blue}{$\hookrightarrow$}\space},
	frame=single
}
\begin{lstlisting}[caption={Example XML file from \emph{LogChunks}}, label=lst:examples, breaklines=true]
<Examples>
  <Log>C/php@php-src/failed/529279089.log</Log>
    <Keywords>ERROR, FAIL, DIFF</Keywords>
    <Category>0</Category>
    <Chunk>001+ ** ERROR: process timed out **

@@

For each repository, \emph{LogChunks} gives about 10 \texttt{Examples}.
Each \texttt{Example} consists of:
\begin{itemize}
	\item \texttt{Log:} the relative path to the input build log.
	\item \texttt{Chunk:} the targeted information chunk. We are targeting the substring of the log that describes why the build failed.
	\item \texttt{Keywords:} keywords a developer would use to search for the log chunk.
	\item \texttt{Category:} a categorization of the structural representation of the log chunk within the build log.
				The category is relative to the other examples for the same repository.
\end{itemize}
Following, this section defines in more detail the labeled log chunk, search keywords and structural categories.

@@

\paragraph{Keywords}
The \texttt{Keywords} contain a list of one to three keywords appearing within the \texttt{Chunk} or in the area around it in the build log.
We aim to select keywords a developer would use to search for the \texttt{Chunk} when inspecting the build log.

@@

>>>>

\chapter{LogChunks Data Set}
\label{sec:data-set}
\includegraphics[page=5, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
This chapter describes the creation of the \emph{LogChunks} data set that we use in our empirical comparison of chunk retrieval techniques.
\emph{LogChunks} is a collection of 797 Travis CI build logs from 80 GitHub repositories spread over 29 programming languages.
For each build log we manually label which substring describes why the build failed.
The data set also provides keywords a developer would use to search for the labeled log chunk and categorizes the log chunks according to their format within the log.

@@

\section{Motivation}
We create \emph{LogChunks} as a data set for our empirical comparison study of the three chunk retrieval techniques PBE, CTS and KWS\@.
In the following, we motivate which repositories and build logs we collect and what data we label for each build log.

\paragraph{Log Collection}
Existing build log analysis approaches, like writing a custom parser and creating regular expressions, are specific to the build tools which create the targeted log output.
In contrast to that, our findings about the chunk retrieval techniques should not be specific to a specific build tool.
As the main development language in a repository influences the build tools, we want our comparison study and \emph{LogChunks} cover a wide array of development languages.
This is why we selected GitHub repositories from 29 different main development languages for our data set.

Examples configure the chunk retrieval techniques we investigate.
As the structure of a build log changes from project to project, these examples are specific to one project.
A user of a chunk retrieval technique has to manually create the examples for for each targeted information and project.
If the effort of creating examples should not too high, e.g. as high as developing a custom extraction program.
For that reason, we choose ten logs and examples per repository for \emph{LogChunks}.

\paragraph{Targeted Log Chunk}
To evaluate chunk retrieval techniques, we have to select an information chunk the techniques should target.
The targeted information chunk should occur as a realistic use case for developers or researchers and should not be trivial for any of the studied techniques.

As the targeted log chunk we choose the part of the log describing why the build failed.
Developers examine build logs to identify the cause of a build failure so they can address the issue.
As we described in Section \ref{sec:rw-ci}, various researchers analyze build logs for the reason the build failed.

In addition, the log chunk describing why a build failed can be represented in different ways in build logs.
The representation and content of this log chunk depends on the tool reporting the fault and a build can fail in various of the tools involved in the build.
This makes it non-trivial for any of the three techniques to extract.
An example of a log chunk easily extractable by all of the three techniques is the \emph{TravisTiming}, described in Section \ref{sec:bli}.
Travis CI logs its timing statements in a always consistent structure, therefore it would be easy for PBE to synthesize a regular expression program to extract them.
The same holds for search keywords identifying the lines which contain the \emph{TravisTiming} and selecting lines with similar text content.

\paragraph{Additional Data: Keywords and Structural Categories}
For each log chunk example we additionally annotated keywords a developer would use to search for this log chunk, as well as a categorization of the chunks into structural categories according to their format within the log.
The technique KWS is configured by keywords instead of in/output examples like the other techniques.
To make the evaluation more consistent an comparable we choose to annotate keywords to each log chunk example.
We can then use these extended examples as a configuration for KWS.
An increasing number of examples is then comparable to the user basing his given keywords on more previously seen build logs.

For our second research question we want to investigate how structurally similar the training examples have to be for a chunk retrieval technique to be applicable.
To measure this in our study we choose to categorize the examples according to their format within the log.
Section \ref{sec:data-schema} gives examples of log chunks from the same and different structural categories.
As we evaluate the techniques separately for separate projects, the categorization is relative to the project which produced the build log.

@@

\lstset{
  language=XML,
  morekeywords={Examples, Example, Log, Keywords, Category, Chunk},
  postbreak=\mbox{\textcolor{cyan}{$\hookrightarrow$}\space},
	frame=single,
	escapeinside=**
}
\begin{figure}[]
	\centering
\begin{lstlisting}[breaklines=true]
<Examples>
  <Log>C/php@php-src/failed/529279089.log</Log>
    <Keywords>ERROR, FAIL, DIFF</Keywords>
    <Category>0</Category>
    <Chunk>001+ ** ERROR: process timed out **

@@

	\caption{Example XML file from \emph{LogChunks}}
	\label{lst:examples}
\end{figure}

For each repository, \emph{LogChunks} gives about 10 \texttt{Examples}.
Each \texttt{Example} consists of:
\begin{itemize}
	\item \textbf{\texttt{Log:}} the relative path to the input build log.
	\item \textbf{\texttt{Chunk:}} the targeted information chunk. We are targeting the substring of the log that describes why the build failed.
	\item \textbf{\texttt{Keywords:}} keywords a developer would use to search for the log chunk.
	\item \textbf{\texttt{Category:}} a categorization of the structural representation of the log chunk within the build log.
				The category is relative to the other examples for the same repository.
\end{itemize}
Following, this section defines in more detail the labeled log chunk, search keywords and structural categories.

@@

\paragraph{Keywords}
The \texttt{Keywords} contain a list of one to three strings appearing within the \texttt{Chunk} or in the area around it in the build log.
We aim to select keywords a developer would use to search for the \texttt{Chunk} when inspecting the build log.
Some example keywords from \emph{LogChunks} are: ``\texttt{FAIL}'', ``\texttt{Error}'', ``\texttt{failed}'', ``\texttt{[-]}'', ``\texttt{===================================}'' or ``\texttt{ERR!}''.

@@

\begin{figure}[]
	\centering
\begin{lstlisting}[breaklines=true]
[0K$ ./sapi/cli/php run-tests.php -P -d extension=`pwd`/modules/zend_test.so $(if [ $ENABLE_DEBUG == 0 ]; then ...
*{\color{cyan}-=-=-=-=-=-}*
Illegal switch 'j' specified!
*{\color{cyan}-=-=-=-=-=-}*
Synopsis:
\end{lstlisting}

	\caption{Log chunk with one line of context from a \emph{different} structural category than the log chunk presented in Listing \ref{lst:examples}, ``{\color{cyan}-=-=-=-=-=-}'' separates the log chunk from the context}
	\label{lst:same-category}
\end{figure}

\begin{figure}[]
	\centering
\begin{lstlisting}[breaklines=true]
========DIFF========
*{\color{cyan}-=-=-=-=-=-}*
005+     Parameter #1 [ <optional> $flags ]

005-     Parameter #1 [ <optional> $ar_flags ]
========DONE========
FAIL Bug#71412 ArrayIterator reflection parameter info [ext/spl/tests/bug71412.phpt] 
*{\color{cyan}-=-=-=-=-=-}*
TEST 9895/13942 [2/2 concurrent test workers running]
\end{lstlisting}

	\caption{Log chunk with one line of context from the \emph{same} structural category than the log chunk presented in Listing \ref{lst:examples}, ``{\color{cyan}-=-=-=-=-=-}'' separates the log chunk from the context}
	\label{lst:different-category}
\end{figure}

### 2019-12-13 thesis/txt/5-study.tex
adding various examples to rw, data set and study

>>>>

\lstset{
  morekeywords={Test Output, Desired Test Output},
  postbreak=\mbox{\textcolor{cyan}{$\hookrightarrow$}\space},
  showspaces=false,
  showstringspaces=false,
  keywordstyle=\color{blue},
	frame=single,
  extendedchars=false,
  texcl=false
}
\begin{figure}[]
  \centering
  \begin{lstlisting}[breaklines=true]
Test Output:
Error: Invalid CSS after "2.3em": expected expression (e.g. 1px, bold), was ";"
        on line 86 of sass/components/dropdown.sass

Desired Test Output:
Error: Invalid CSS after "2.3em": expected expression (e.g. 1px, bold), was ";"
        on line 86 of sass/components/dropdown.sass
        from line 5 of sass/components/_all.sass
        from line 6 of bulma.sass
  \end{lstlisting}  
  \caption{Example for an unsuccessful retrieval, PBE retrieved only two of the four targeted lines}
  \label{lst:system-log}
\end{figure}

### 2019-12-13 thesis/txt/6-discussion.tex
adding various examples to rw, data set and study

\subsection{Program Synthesis by Example (PBE)}
\paragraph{Configuration and Input}
Our study results show that chunk retrieval with PBE gives best results when the training examples are from one structural category.
This means it is suited to retrieve information chunks that always have the same surrounding structure.
To extract for example the reason a build failed, the log passage describing the failure would always have to be started and ended the same way.

@@

\section{Recommendations of Suitable Techniques}
% This section unifies our study results of the chunk retrieval techniques PBE, CTS and KWS\@.
% It recommends which technique to use depending on the available configuration and 
After discussing three chunk retrieval techniques separately we now want to unify out results into one recommendation scheme.
We present a decision tree, which asks questions to a developer or researcher who wants to retrieve information chunks from build logs.
The answers lead the user to more questions or the recommended technique as leaf nodes.
Figure~\ref{fig:crt-recommendation} presents the decision tree.

@@

\section{Threats to Validity}

There are several threats to the validity of the conclusions of our work.

>>>>

\subsection{Program Synthesis by Example (PBE)}

\paragraph{Configuration and Input}
Our study results show that chunk retrieval with PBE gives best results when the training examples are from one structural category.
This means it is suited to retrieve information chunks that always have the same surrounding structure.
To extract for example the reason a build failed, the log passage describing the failure would always have to be started and ended the same way.

@@

\section{Recommendations of Suitable Techniques}
After discussing three chunk retrieval techniques separately we now want to unify out results into one recommendation scheme.
We present a decision tree, which asks questions to a developer or researcher who wants to retrieve information chunks from build logs.
The answers lead the user to more questions or the recommended technique as leaf nodes.
Figure~\ref{fig:crt-recommendation} presents the decision tree.

@@

\section{Threats to Validity}
There are several threats to the validity of the conclusions of our work.
